Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-02-11 docs/00-manuscripts/ieee/ieee-manuscript.tex
update rw4 replies

Inspecting the results of the \ac{PCA} for all feature sets, HR showed a high correlation among variables with a \ac{POV} of 84\% in the first \ac{PC} and 14\% in the second \ac{PC}, summing up to a cumulative \ac{POV} of 98\% (\autoref{fig:pca-pov}).

Feature sets VI (51\% PC1, 25\% PC2, 7\% PC3) and NRI (51\% PC1, 25\% PC2, 8\% PC3) showed a more diverse distribution across their respective \ac{PC}s.

@@

@@

\subsection{Data quality}

Environmental datasets always come with some constraints that can have potential influence on the modeling process and its outcome.
The following paragraph discusses these.
The decision to use a buffer of two meters for the extraction of reflectance values was a complex process.
Due to the reported geometric offset of up to 1 m within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different observation was reasonably high.
By using a buffer of two meters (or more), the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
However, when using no buffer at all, the difference between single, neighboring pixels might even be higher compared to a smoothed value composed out of a buffer.
It was concluded that using a buffer of two meters is a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.

@@

% parameter limits
\begin{table}[h!]
	\centering
	\caption[t]{Hyperparameter ranges and types for each model.
		Hyperparameter notations from the respective R packages were used.}
	\begingroup\scriptsize
	\begin{tabularx}{.5\textwidth}{llXXrXX}
		\\
		\specialcell{Model                                              \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
		\toprule
		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
		 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
		\midrule
		\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
		 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
		\midrule
		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
		 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
		 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
		 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
		 & \texttt{gamma}              & int & 0        & 10      & 0   \\
		 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
		 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
		\bottomrule
	\end{tabularx}
	\endgroup
	\label{tab:hyperparameter_limits}
\end{table}

>>>>

The PCA-based correlation analysis in \autoref{fig:pca-pov} shows that, depending on the feature set, between two (HR) and five (VI) \ac{PC} are needed to explain 90\% of variance.
The less PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
Hence, a high intra-feature correlation can be observed for feature set HR which reaches 98\% with only two PCs.
Feature sets NRI and VI are more diverse and require four (NRI) and five (VI) PCs to explain 90\% of the variance in the data, respectively.

@@

In addition, data quality issues (\autoref{subsec:data-quality}) might have an influence on model performances.
These include the timing of the acquisition of the hyperspectral data (late phenological phase), field measurement errors when surveying defoliation, the influence of background reflectance (e.g. soil reflectance) and the possible geometrical offset of trees in the field.

@@

\subsection{Data Quality}
\label{subsec:data-quality}

Environmental datasets always come with some constraints that can have potential influence on the modeling process and its outcome.
The following paragraph discusses these.

Finding a suitable approach to extract the remote sensing information from each tree was a complex process.
Due to the reported geometric offset of up to 1 m within the hyperspectral data, the risk of assigning a value to an observation which would actually refer to a different observation was reasonably high.
Finally it was decided to apply a buffer around each recorded tree location and not to solely rely on the reported location during value extraction.
By using a buffer of multiple meters, the probability of including information from other trees into the final value increases, blurring the actual value of the tree observation.
However, when using no buffer at all, the difference between single, neighboring pixels might be quite high and involves the risk of including non-tree data.
It was concluded that using a buffer of two meters can be a good compromise between the inclusion of information from too many surrounding trees and not accounting for the geometric offset at all.
Even though no results showing the influence of multiple buffer values on the extraction were provided, it is hypothesized that the relationships between features would not change substantially, leading to almost identical model results.
Besides using a buffer to extract the hyperspectral information, a segmentation could be considered.
However, this method would require more effort for no clear added value in our view and move the focus of this manuscript more to data preprocessing and away from the focus on filter-based feature selection methods.

@@

% parameter limits
\begin{table}[h!]
	\centering
	\caption[t]{Hyperparameter ranges and types for each model.
		Hyperparameter notations from the respective R packages were used.}
	\begingroup\scriptsize
	\begin{tabularx}{.5\textwidth}{llXXrXX}
		\\
		\specialcell{Model                                              \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
		\toprule
		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_{try}$}    & dbl & 0         & 0.5      & -       \\
		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
		 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
		\midrule
		\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
		 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
		\midrule
		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
		 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
		 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
		 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
		 & \texttt{gamma}              & int & 0        & 10      & 0   \\
		 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
		 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
		\bottomrule
	\end{tabularx}
	\endgroup
	\label{tab:hyperparameter_limits}
\end{table}
